% Paper 8 of The Windstorm Series
\documentclass[12pt,a4paper]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{booktabs}

\title{The Vision Basin: Cross-Modal Throughput Measurement Reveals Modality-Specific Information Extraction Rates}

\author{Grant Lavell Whitmer III\\
The Windstorm Institute, Fort Ann, NY 12827, USA\\
\texttt{grantwhitmer3@gmail.com}}

\date{April 2026}

\begin{document}

\maketitle

\begin{abstract}
Paper 7 established that the language-model throughput basin reflects training-data entropy minus exploitable hierarchical structure: $\mathrm{BPT} \approx \mathrm{source\_entropy} - f(\mathrm{structural\_depth})$. If this equation is general, different modalities should sit at different throughput levels. We test this prediction across three modalities --- language, vision, and audio --- using twelve pretrained models, from-scratch training, source-entropy measurement via lossless compression, and structural-destruction cascades. Each modality has its own characteristic throughput: language at 0.85--1.30 bits per source byte (nine models, two corpora), generative vision at 1.33--1.36 bits per pixel (MAE reconstruction), and real speech at 1.89 bits per mel-spectrogram dimension (LJ Speech, 3{,}000 utterances, three seeds). A visual shuffling cascade on STL-10 reveals a spatial structural bonus of 0.69 bits per pixel. Music extracts 2.69 bits per mel dimension. Bits per pixel is not patch-size-invariant (varies $4\times$ across patch sizes), and visual source entropy is resolution-dependent. Version 2.1 added a from-scratch verification round: a 112-million-parameter ViT-MAE trained at seven controlled-entropy image distributions with three seeds each (Welch $t = 249{,}994$, $p = 1.6 \times 10^{-11}$, Cohen's $d = 204{,}119$). Version 2.2 (April 2026) adds the publication-readiness presentation: a methodological-journey narrative ($\S$1.3) and an adversarial-review-defense table ($\S$4.4); Zenodo \texttt{10.5281/zenodo.19672828} (concept DOI \texttt{10.5281/zenodo.19672827}). This is the eighth paper in The Windstorm Series.
\end{abstract}

\noindent\textbf{Keywords:} vision basin; audio throughput; cross-modal comparison; masked autoencoder; source entropy; structural bonus; bits per source unit.

\section{Introduction}

The full text of this paper is provided as \texttt{paper.pdf} (current version v2.2) in this repository. The markdown source for the latest version is in \texttt{paper/Paper8-v2.2-draft.md} in this repository (with historical \texttt{paper/Paper8-v2.1-draft.md} alongside).

\section*{Code and Data Availability}

All code, experimental data, and figures are available under MIT license at \url{https://github.com/Windstorm-Institute/vision-basin} and \url{https://github.com/Windstorm-Labs/vision-basin}. The Grand Slam Supplementary Materials providing statistical verification (v2.1) and the v2.2 methodological-journey + adversarial-review-defense framework are at \texttt{grandslam\_supplementary.pdf} and \texttt{paper.pdf} (current version v2.2) in this repository.

\end{document}
